---
title: "Predicting Lesions with FLEXCONN"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Predicting Lesions with FLEXCONN}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(neurobase)
library(flexconn)
```


## Installing FLEXCONN

FLEXCONN stands for Fast Lesion Extraction using Convolutional Neural Networks, which the main goal is to segment lesions from magnetic resonance images (MRIs) from patients with multiple sclerosis (MS).  The methods are described here:
https://arxiv.org/abs/1803.09172.  Patients with MS commonly have 2 MRI sequences done, T1-weighted images and fluid attenuated inversion recovery (FLAIR).  FLEXCONN requires a number of Python modules to run.  You can install them either system wide (not shown) or using `reticulate`.  The `reticulate` way is wrapped in `install_flexconn`:

```{r, eval = FALSE}
flexconn::install_flexconn()
```

## Downloading a Model 

Here we download one of the fit models from FLEXCONN 1.1 (https://www.nitrc.org/projects/flexconn/), which is from one reader, a `35x35` patch size (2D model) and it was fit using 21 atlases.  

```{r dl_model}
model_file = tempfile(fileext = ".h5")
base_url = paste0(
  "https://github.com/muschellij2/flexconnr",
  "/raw/master/inst/extdata/")
model_url = paste0(
  base_url, "21atlases/",
  "atlas_with_mask1/FLEXCONNmodel2D_35x35_17-10-2017_21-53-35.h5")
download.file(model_url, destfile = model_file)
```

We will load the model into `R` so we can predict using it
```{r load_model}
model = keras::load_model_hdf5(model_file)
model
```

## Downloading Data

Again, we need a T1-weighted and FLAIR image to do prediction.  We use some example data here from the package and plot each one.  The `gt` object is the ground truth, manual segmentation:

```{r dl_data}
t1_file = system.file("extdata", "MPRAGE.nii.gz", package = "flexconn")
t1 = readnii(t1_file)
ortho2(robust_window(t1))

flair_file = system.file("extdata", "FLAIR.nii.gz", package = "flexconn")
flair = readnii(flair_file)
ortho2(robust_window(flair))

gt_file = system.file("extdata", "MPRAGE_manual.nii.gz", package = "flexconn")
gt = readnii(gt_file)
ortho2(robust_window(flair), gt, xyz = xyz(gt))
```

## Predicting Lesions 

Here we will predict using the pre-trained model and we 
```{r predict_model}
res = flexconn_predict(model,
                       t1 = t1_file,
                       flair = flair_file)
```

Here we overlay the probabilities on the FLAIR image:
```{r plot_result}
col.y =  c("#010000", "#0C0000", "#170000", "#210000", "#2C0000",
            "#360000", "#410000", "#4C0000", "#560000", "#610000",
            "#6C0000", "#760000", "#810000", "#8B0000", "#960000",
            "#A10000", "#AB0000", "#B60000", "#C10000", "#CB0000",
            "#D60000", "#E00000", "#EB0000", "#F60000", "#FF0100",
            "#FF0C00", "#FF1700", "#FF2100", "#FF2C00", "#FF3600",
            "#FF4100", "#FF4C00", "#FF5600", "#FF6100", "#FF6C00",
            "#FF7600", "#FF8100", "#FF8B00", "#FF9600", "#FFA100",
            "#FFAB00", "#FFB600", "#FFC100", "#FFCB00", "#FFD600",
            "#FFE000", "#FFEB00", "#FFF600", "#FFFF02", "#FFFF12",
            "#FFFF22", "#FFFF32", "#FFFF42", "#FFFF52", "#FFFF62",
            "#FFFF72", "#FFFF81", "#FFFF91", "#FFFFA1", "#FFFFB1",
            "#FFFFC1", "#FFFFD1", "#FFFFE1", "#FFFFF1")
col.y = grDevices::col2rgb(col.y, alpha = 0.5)
col.y = grDevices::rgb(col.y[1,], col.y[2,], col.y[3,], col.y[4,], maxColorValue = 255)
ortho2(robust_window(flair), res, col.y = col.y)
```

Here we see we have a probability map of lesions.  To get a binary mask, we can use the `remove_small_components`, which thresholds the image and removes voxels with fewer than 27 connected voxels:

```{r make_binary}
pred = remove_small_components(res)
```

```{r plot_binary}
ortho2(robust_window(flair), pred, xyz = xyz(pred), col.y = "#FF000080")
```

```{r dice}
fast_dice(pred, gt)
```
